---
title: "Task 5"
author: Tereza Kozurkova
date: "10/14/2024"
format:
  html:
    embed-resources: true
editor: source
---
## Task 5: Visualization

Using the dataset from **Task 4.2** (`airbnb.csv`), create **four visualizations** using
both the **base plotting system** and **ggplot2**. You may use arbitrary subsets of
the data for this task. Ensure that all visualizations include:\
- A title,\
- Annotated axes,\

- A legend (if needed).

### Instructions:

1. **Histogram**:\
Create a histogram using the `hist` function from the **base plotting system**.

2. **Scatter Plot**:\
Create a scatter plot using **ggplot2** with the `geom_point` geometry. Incorporate
a third variable by using different colors to represent it.

3. **Boxplot**:\
Create a boxplot using **ggplot2** with the `geom_boxplot` geometry.

4. **Custom Plot**:\
Create an additional plot of your choice using either the **base plotting system** or
**ggplot2**.

```{r}
# Cesta k priečinku, kde sa nachádzajú .csv súbory
folder_path <- "C:/aibnb_data"

# Funkcia na zlúčenie a uloženie datasetu
prepare_dataset <- function(folder_path) {
  # Načítanie všetkých súborov v priečinku
  files <- list.files(folder_path, full.names = TRUE)
  
  # Inicializácia prázdneho datasetu
  merged_data <- data.frame()
  
  for (file_path in files) {
    # Extrakcia názvu mesta a typu dňa
    file_name <- basename(file_path)
    parts <- strsplit(file_name, "_")[[1]]
    city_name <- parts[1]
    day_type <- gsub("\\.csv$", "", parts[2])
    
    # Načítanie dát
    data <- read.csv(file_path, header = TRUE)
    
    # Pridanie nových stĺpcov
    data$city <- city_name
    data$day_type <- day_type
    
    # Zlúčenie dát
    merged_data <- rbind(merged_data, data)
  }
  
  # Uloženie výsledného datasetu
  write.csv(merged_data, file = file.path(folder_path, "airbnb.csv"), row.names = FALSE)
  return(merged_data)
}

# Spustenie funkcie
airbnb_data <- prepare_dataset(folder_path)
```

```{r}
# Načítanie datasetu z Task 4.2
airbnb_data <- read.csv("C:/aibnb_data/airbnb.csv", header = TRUE)

# Histogram cien (realSum) - Base R
hist(
airbnb_data$realSum,
breaks = 50,
col = "skyblue",
main = "Histogram of Airbnb Prices",
xlab = "Price (realSum)",
ylab = "Frequency"
)
```

```{r}
# Scatter plot - Base R
plot(
airbnb_data$guest_satisfaction_overall,
airbnb_data$realSum,
main = "Scatter Plot of Prices vs Guest Satisfaction", # Titul
xlab = "Guest Satisfaction (overall)", # Popis osi X
ylab = "Price (realSum)", # Popis osi Y
col = as.factor(airbnb_data$city), # Rôzne farby podľa mesta
pch = 16, # Typ bodu
cex = 0.5 # Veľkosť bodov
)

# Pridanie legendy
legend("topright", legend = levels(as.factor(airbnb_data$city)),
col = 1:length(levels(as.factor(airbnb_data$city))),
pch = 16, title = "City")
```

```{r}
# Boxplot cien podľa mesta - Base R
boxplot(
realSum ~ city,
data = airbnb_data,
main = "Boxplot of Airbnb Prices by City", # Titul
xlab = "City", # Popis osi X
ylab = "Price (realSum)", # Popis osi Y
col = "lightblue", # Farba boxplotu
border = "darkblue" # Farba okrajov
)

```

```{r}
# Filtrácia dát na pracovné dni a víkendy
weekend_data <- airbnb_data[airbnb_data$day_type == "weekends", ]
weekday_data <- airbnb_data[airbnb_data$day_type == "weekdays", ]

# Hustotný graf - Base R
plot(density(weekend_data$realSum),
main = "Density Plot of Airbnb Prices (Weekdays vs Weekends)", # Titul
xlab = "Price (realSum)", # Popis osi X
ylab = "Density", # Popis osi Y
col = "blue", # Farba pre víkendy
lwd = 2) # Hrúbka čiary
lines(density(weekday_data$realSum), col = "red", lwd = 2) # Červená čiara pre pracovné dni

# Pridanie legendy
legend("topright", legend = c("Weekends", "Weekdays"), col = c("blue", "red"), lwd = 2)
```



